Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework

### 2014-08-29 tex/manuscript.tex
Additional edits, as well as added first draft of proposal letter

\usepackage{natbib}

@@

\usepackage{authblk} % adds some nice options for displaying the author list

@@

% available colors:
% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
%
\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}

@@

% uncomment to disable individual authors
% \renewcommand{\mt}[1]{\relax}%
% \renewcommand{\ib}[1]{\relax}%
% \renewcommand{\aa}[1]{\relax}%
% \renewcommand{\ia}[1]{\relax}%
% \renewcommand{\db}[1]{\relax}%
% \renewcommand{\ab}[1]{\relax}%
% \renewcommand{\rd}[1]{\relax}%
% \renewcommand{\mjf}[1]{\relax}%
% \renewcommand{\tf}[1]{\relax}%
% \renewcommand{\dm}[1]{\relax}%
% \renewcommand{\ks}[1]{\relax}%
% \renewcommand{\ns}[1]{\relax}%
% \renewcommand{\wt}[1]{\relax}%
% \renewcommand{\dg}[1]{\relax}%

@@

\title{Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework}
\date{}
\author[1,2]{Matthew V. Talluto}
\author[1,2]{Isabelle Boulangeat}
\author[3]{Aitor Ameztegui}
\author[4]{Isabelle Aubin}
\author[1,2,5]{Dominique Berteaux}
\author[1,2]{Alyssa Butler}
\author[6,7]{Frédérik Doyon} %??
\author[8]{C. Ronnie Drever}
\author[9]{Marie-Josée Fortin}
\author[1]{Tony Franceschini}
\author[10]{Jean Liénard}
\author[4]{Dan McKenney}
\author[2,3]{Kevin A. Solarik}
\author[10]{Nickolay Strigul}
\author[11, 12]{Wilfried Thuiller}
\author[1,2]{Dominique Gravel}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[6]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
\affil[7]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
\affil[8]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
\affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
\affil[10]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
\affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}

@@

\begin{document}

%% ----------------------------------
%
%     TITLE PAGE
%
%% ----------------------------------

\maketitle

\begin{flushleft}
\textbf{Keywords:} Uncertainty, scaling, range dynamics, trees, disturbances, patterns and processes, predict or understand, climate change, biotic interactions, spatial ecology, decision making, metamodeling, species distribution modeling
\end{flushleft}

\begin{abstract}
There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
However, in general, individual approaches include only a small subset of the total available information about a species. 
For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
Furthermore an inherent problem with the diversity of approaches is that different models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
The resulting meta-model framework produces probabilistic estimates of species presence with uncertainty that reflects error from data sources as well process error from all sub-models.
 These predictions reflect all of the information used as input for the original scale-specific sub-models.
We illustrate the metamodeling approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
Although this hierarchical Baysian method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
\end{abstract}

@@

>>>>

\usepackage{setspace}
\usepackage{natbib}

@@

\usepackage{authblk} % adds some nice options for displaying the author list
\renewcommand\Authsep{\protect\\}
\renewcommand\Authands{\protect\\}

@@

% available colors:
% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
%
\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}

@@

% uncomment to disable individual authors
% \renewcommand{\mt}[1]{\relax}%
% \renewcommand{\ib}[1]{\relax}%
% \renewcommand{\aa}[1]{\relax}%
% \renewcommand{\ia}[1]{\relax}%
% \renewcommand{\db}[1]{\relax}%
% \renewcommand{\ab}[1]{\relax}%
% \renewcommand{\fd}[1]{\relax}%
% \renewcommand{\rd}[1]{\relax}%
% \renewcommand{\mjf}[1]{\relax}%
% \renewcommand{\tf}[1]{\relax}%
% \renewcommand{\dm}[1]{\relax}%
% \renewcommand{\ks}[1]{\relax}%
% \renewcommand{\ns}[1]{\relax}%
% \renewcommand{\wt}[1]{\relax}%
% \renewcommand{\dg}[1]{\relax}%

@@

\title{Cross-scale integration of knowledge for predicting species ranges: a metamodeling framework}
\date{}
\author[1,2,13]{Matthew V. Talluto (mtalluto@gmail.com)}
\author[1,2]{Isabelle Boulangeat}
\author[3]{Aitor Ameztegui}
\author[4]{Isabelle Aubin}
\author[1,2,5]{Dominique Berteaux}
\author[1,2]{Alyssa Butler}
\author[6,7]{Frédérik Doyon}
\author[8]{C. Ronnie Drever}
\author[9]{Marie-Josée Fortin}
\author[1]{Tony Franceschini}
\author[10]{Jean Liénard}
\author[4]{Dan McKenney}
\author[2,3]{Kevin A. Solarik}
\author[10]{Nickolay Strigul}
\author[11, 12]{Wilfried Thuiller}
\author[1,2]{Dominique Gravel}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[6]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
\affil[7]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
\affil[8]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
\affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
\affil[10]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
\affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[13]{Author for correspondance. Address: Departament de Biologie, chimie, et geographie, 300, Allée des Ursulines, Rimouski, Quebec G5L 3A1, Canada}

@@

\begin{document}
\doublespacing
%% ----------------------------------
%
%     TITLE PAGE
%
%% ----------------------------------

\begin{titlingpage}
	\maketitle

	\mt{need to include email addresses of all authors}

	\begin{flushleft}
	\textbf{Short title: } \mt{less than 45 chars including spaces}

	\textbf{Type of article: } Ideas and Perspectives

	\textbf{Word counts: } \mt{main text excludes abstract, acknowledgements, references, table and figure legends; limited to 7500 words}
	\begin{itemize}
		\item[] Abstract: 
		\item[] Main Text: 
		\item[] Boxes:
	\end{itemize}

	\textbf{Number of references: } \mt{limited to 80}

	\textbf{Number of figures: } 6

	\textbf{Number of tables: } 0

	\textbf{Number of text boxes: } 1

	\textbf{Keywords:} Climate change, decision making, metamodeling, patterns and processes, range dynamics, scaling, spatial ecology, species distribution modeling, uncertainty
	\end{flushleft}
\end{titlingpage}

\begin{abstract}
\mt{limited to 200 words, currently at 235}
There is great interest in modeling species ranges, resulting in a great diversity of modeling approaches.
However, in general, individual approaches include only a small subset of the total available information about a species. 
For example, a broad-scale correlative model might use climate variables to predict presence or absence, but ignore what is known about smaller-scale processes such as the effect of local climate on growth, fecundity, and dispersal rates.
Furthermore an inherent problem with the diversity of approaches is that different models often produce distinct predictions for the same species, with no simple way to reconcile these predictions.
We present a flexible framework for integrating models at multiple scales using hierarchical Bayesian methods. 
The resulting meta-model framework produces probabilistic estimates of species presence with uncertainty that reflects error from data sources as well process error from all sub-models.
 These predictions reflect all of the information used as input for the original scale-specific sub-models.
We illustrate the metamodeling approach using two examples, and demonstrate that the framework can substantially reduce uncertainty when projecting beyond the range of some of the original data sources.
We conclude by discussing the application of our method and its accessibility to conservation biologists and land managers.
Although this hierarchical Baysian method requires extensive statistical programming experience, we anticipate the results will be of wide application and interest, and we therefore encourage collaboration between modelers and practitioners in applying our framework.
\end{abstract}

@@

\input{figures}

%% ----------------------------------
%
%     TEXT BOX
%
%% ----------------------------------
